LEMONADE: Leveraging Modern Automated Deduction for Program Synthesis\\ \vspace{0.5cm} \large{WWTF ICT 2022}

### 2022-03-31 WWTF/main.tex
asdf

The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers \sout{(}and SMT solvers\sout{???)}, which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).

@@

Much of the theoretical groundwork for implementing such a system is already present,
but is scattered and mostly limited to theoretical settings.
We sketch here the two main directions that connect the Curry-Howard correspondence to classical logic:
1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation.
However in this process we must in general give up on some desirable properties, at the very least the realizability of $\vee$ and $\exists$.
2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g. proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA}, {\color{blue}this point is of particular interest for SMT}.
We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
A) We will develop the first\emph{ program extraction calculi} for \emph{resolution resp. superposition proofs}.
In particular, we need to pay careful attentiont to the normal-form transformation --- a pre-processing step --- in automated theorem provers, which also involve equivalences that are only classically valid.
B) We will identify assumptions under which our adaptions of the techniques 1) and 2) guarantee that such a program extraction is possible.
We are not aware of related results in the literature.

@@

2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}. There have been many variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
We will collect, adapt and extend these results with regard to our concrete setting. 
This is particularly promising when  {\color{blue} applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when} considering altered specifications (see next point).
%We expect that in practice these approaches will be sufficient in many situations.

>>>>

The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers (including SMT solvers), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).

@@

Much of the theoretical groundwork for implementing such a system is already present,
but is scattered and mostly limited to theoretical settings.
We sketch here the two main directions that connect the Curry-Howard correspondence to classical logic:
1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation.
However in this process we must in general give up on some desirable properties, at the very least the realizability of $\vee$ and $\exists$.
2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g. proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA} (this point is of particular interest for SMT).
We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
A) We will develop the first\emph{ program extraction calculi} for \emph{resolution resp. superposition proofs}.
In particular, we need to pay careful attentiont to the normal-form transformation --- a pre-processing step --- in automated theorem provers, which also involve equivalences that are only classically valid.
B) We will identify assumptions under which our adaptions of the techniques 1) and 2) guarantee that such a program extraction is possible.
We are not aware of related results in the literature.

@@

2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}. There have been many variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
We will collect, adapt and extend these results with regard to our concrete setting.
This is particularly promising when applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when considering altered specifications (see next point).
%We expect that in practice these approaches will be sufficient in many situations.
